Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-05 latex/robdml.tex
tex
svg

@@

\begin{figure}
	\oo{[diagram]}
	\caption{\oo{Diagram demonstrating the core idea}}
\end{figure}

@@

In Deep Metric Learning (DML)~\cite{revisiting,dmlreality}, an embedding
function $\phi:\mathcal{X}\mapsto \Phi \subseteq \mathbb{R}^D$ is learned to
map data points $X\in\mathcal{X}$ into an embedding space $\Phi$, where the
output of $\phi(\cdot)$ is usually normalized to the real unit hypersphere for
regularization purpose.
%
With a predefined distance function $d(\cdot,\cdot)$, it allows us to measure
the distance between $X_i$ and $X_j$ as
$d_\phi(X_i,X_j)=d(\phi(X_i),\phi(X_j))$.
%
Various loss functions for learning the underlying embedding function have been
proposed~\cite{revisiting,dmlreality}, where the triplet loss~\cite{facenet}
remain to be a strong and widely used baseline that could reach
state-of-the-art performance with appropriate triplet sampling strategy.

@@

Although attacks against DML has been widely studied~\cite{advrank,advorder},
the defense methods for improving adversarial robustness is much less explored.
%
As adversarial training~\cite{madry} remains to be one of the most effective
defense for classification, existing defense methods for DML are also adversarial
training-based.

\oo{[Quick review of existing defense methods]}
EST~\cite{advrank} and ACT~\cite{robrank}. Their issues.

>>>>

\begin{figure}[t]
	\includegraphics[width=1.0\columnwidth]{introdiag.pdf}
	\caption{\oo{Performance plot with aligned conditions.
	(1) robustness and r1 under the same training cost;
	(2) r1 and training cost for reaching similar robustness;
	(3) training cost and robustness for reaching similar r1.}}
\end{figure}

@@

@@

In Deep Metric Learning (DML)~\cite{revisiting,dmlreality}, an embedding
function $\phi:\mathcal{X}\mapsto \Phi \subseteq \mathbb{R}^D$ is learned to
map data points $X\in\mathcal{X}$ into an embedding space $\Phi$, where the
output of $\phi(\cdot)$ is usually normalized to the real unit hypersphere for
regularization purpose.
%
With a predefined distance function $d(\cdot,\cdot)$, it allows us to measure
the distance between $X_i$ and $X_j$ as
$d_\phi(X_i,X_j)=d(\phi(X_i),\phi(X_j))$.
%
Various loss functions for learning the underlying embedding function have been
proposed~\cite{revisiting,dmlreality}, where the triplet loss~\cite{facenet}
remain to be a strong and widely used baseline that could reach
state-of-the-art performance with appropriate triplet sampling strategy.
%
Given a triplet of image embeddings (anchor $\va=\phi(X_a)$, positive sample
$\vp=\phi(X_p)$, negative sample $\vn=\phi(X_n)$), the triplet loss is defined
as:
%
$
%
	L_\text{trip}(\va, \vp, \vn; \gamma) = \max(0, d(\va, \vp) - d(\va, \vn) +
	\gamma),
%
$
%
where $\gamma$ is a predefined margin parameter.

@@

Although attacks against DML has been widely studied~\cite{advrank,advorder},
the defense methods for improving adversarial robustness is much less explored.
%
As adversarial training~\cite{madry} remains to be one of the most effective
defense for classification, existing defense methods for DML are also
adversarial training-based.
%
Embedding-Shifted Triplet (EST)~\cite{advrank} adopts adversarial counterpars
of $\va,\vp,\vn$ with maximum embedding move distance off their original
locations, \ie,
$L_\text{EST}=L_\text{trip}(\tilde{\va},\tilde{\vp},\tilde{\vn};\gamma)$ where
$\tilde{\va}=\phi(X+r^*)$, $r^*=\arg\max_{r}d_\phi(X+r, X)$.
%
Anti-Collapse Triplet (ACT)~\cite{robrank} collapses the embedding vectors
of positive and negative sample, and enforces the model to separate them
apart, \ie, $L_\text{ACT}=L_\text{trip}(\va, \overrightarrow{\vp},
\overleftarrow{\vn};\gamma)$, where $[\overrightarrow{\vp},\overleftarrow{\vn}]
=[\phi(X_p+r_p^*), \phi(X_n+r_n^*)]$, and $[r_p^*,r_n^*]=\arg\min_{r_p,r_n}
d_\phi(X_p+r_p, X_n+r_n)$.
%
\oo{Their issues.}
